### 2019-01-28 treestats_paper.tex
Tidy up refs and TS intro.

%%%%%%%%%%
\paragraph{Abstract}
Here we describe how to efficiently compute single-site statistics
from population genetic data
using efficient methods on tree sequences as implemented in msprime.
The method is general enough to compute any currently-defined statistic,
and we use this general framework to suggest a number of new statistics.
\plr{say what method is: sthg about weighting fns}
This formulation also makes it easy to see what statistic of the underlying genealogical trees
each sequence-based statistic is estimating,
and, if underlying trees are also available,
the same methods can be used to quickly compute these quantities.

@@

\plr{need something different here}
Today's vast quantity of whole-genome sequence
makes it possible to confidently measure quantities
between \emph{individuals} that previously had to be roughly estimated
averaging over populations.
For instance, XXX estimated $F_{ST}$ between XX from allozyme sequence,
but now we know that $F_{ST}$ can be viewed as $1 - t_{within}/t_{total}$ \citep{slatkin_fst}
which can therefore be estimated precisely by $1 - \pi_{within}/\pi_{total}$.

@@

The \emph{succinct tree sequence} is a data structure that efficiently encodes
the genealogical trees that describe how a set of individuals are related to each other
at each point along their genome.
The data structure was introduced for coalescent simulation by \citet{kelleher2016efficient},
and refined for forwards simulation by \citet{kelleher2018efficient}.
These genealogical trees change as one looks across the genome
because of the results of recombination in ancestors.
However, nearby genealogical trees tend to share a lot of common structure,
and single genealogical relationships (e.g., individual $x$ inherits from individual $y$)
are often shared across relatively long distances,
which manifest as shared \emph{edges} across many adjacent trees in the tree sequence.
\citet{kelleher2016efficient}
showed how this this redundancy can be used
to both store genome sequence and the associated genealogical relationships extremely compactly,
as well as to compute statistics extremely efficiently.
This paper generalizes the latter strategy
to a much broader class of statistics.

\plr{Review of tree shape stuff in phylogenetics}
a good review of theoretical properties \citet{semple_and_steel}.

>>>>

%%%%%%%%%%
\paragraph{Abstract}
Here we describe how to efficiently compute single-site statistics
from population genetic data using the succinct tree sequence
encoding for correlated genealogies.
The method is general enough to compute any currently-defined statistic,
and we use this general framework to suggest a number of new statistics.
\plr{say what method is: sthg about weighting fns}
This formulation also makes it easy to see what statistic of the underlying genealogical trees
each sequence-based statistic is estimating,
and, if underlying trees are also available,
the same methods can be used to quickly compute these quantities.

@@

\plr{need something different here}
Today's vast quantity of whole-genome sequence
makes it possible to confidently measure quantities
between \emph{individuals} that previously had to be roughly estimated
averaging over populations.
For instance, XXX estimated $F_{ST}$ between XX from allozyme sequence,
% JK: FIXME what's the slatkin ref here?
but now we know that $F_{ST}$ can be viewed as $1 - t_{within}/t_{total}$ \citep{slatkin_fst}
which can therefore be estimated precisely by $1 - \pi_{within}/\pi_{total}$.

@@

The \emph{succinct tree sequence} (or tree sequence, for brevity) is a data
structure that efficiently encodes the genealogical trees describing how a set of
individuals are related to each other at each point along their genome. The
data structure was introduced in the context of coalescent
simulation~\citep{kelleher2016efficient}, and lead to scalability increases of
several orders of magnitude over existing methods. It was subsequently extended
and refined for forwards-time simulations~\citep{kelleher2018efficient,haller2018tree}
where similar efficiency gains were made. Recent
work~\citep{kelleher2018inferring} has shown that tree sequence algorithms can
also be used to massively increase the scalability of methods for inferring
genome-wide genealogies, and shown that it is feasible to infer trees for
millions of whole genome samples.

The key to the remarkable scalability and efficiency of tree sequence
algorithms is the way in which the shared structure in adjacent trees along
the genome is encoded. As one looks across the genome the genealogies change
because of the results of recombination in ancestors.
However, nearby trees tend to share a lot of common structure,
and single genealogical relationships (e.g., individual $x$ inherits from individual $y$)
are often shared across relatively long distances,
which manifest as shared \emph{edges} across many adjacent trees in the tree sequence.
This redundancy can be used
to both store genome sequence and the associated genealogical relationships extremely compactly,
as well as to compute statistics highly efficiently~\citet{kelleher2016efficient}.
This paper generalizes the latter strategy to a much broader class of statistics.

\plr{Review of tree shape stuff in phylogenetics}
a good review of theoretical properties \citet{semple2003phylogenetics}.
